Concurrent Programming with CXL

### 2020-06-16 book/paper.tex
Added concept of race condition

\newglossaryentry{atomicity}
{
  name=atomicity,
  description={describes that a certain machine instruction or sequence of machine
    instructions by a process is executed indivisibly and cannot be interleaved with 
    machine instructions of another process}
}
\newglossaryentry{concurrent execution}
{
  name=concurrent execution,
  description={(aka parallel execution) is when there are multiple processes executing and
      their machine instructions are interleaved in an unpredictable manner}
}
\newglossaryentry{context}
{
  name=context,
  description={(aka continuation) describes the state of a running process,
    including its program counter, the values of its variables (stored in its
    register), and the contents of its stack}
}
\newglossaryentry{critical section}
{
  name=critical section,
  description={(aka critical region) is a set of instructions that only one process
    is allowed to execute at a time.  The instructions are, however, not executed
    atomically, as other processes can continue to execute and access shared
    variables}
}
\newglossaryentry{determinism}
{
  name=determinism,
  description={is when the outcome of an execution is uniquely determined by the
    initial state}
}
\newglossaryentry{invariant}
{
  name=invariant,
  description={is a binary predicate over states that must hold for every
  reachable state of a process}
}
\newglossaryentry{machine instruction}
{
  name=machine instruction,
  description={is an atomic operation on the CXL machine, executed by a process}
}
\newglossaryentry{macro step}
{
  name=macro step,
  description={is a sequence of micro steps with at most one instruction
        externally visible to other processes.  It is the unit of interleaving
        in CXL}
}
\newglossaryentry{micro step}
{
  name=micro step,
  description={is the execution of one machine instruction by a process, updating
        its state and the state of the CXL machine}
}
\newglossaryentry{model checking}
{
  name=model checking,
  description={is a formal verification method that explores all possible executions
    of a program, which must have a finite number of states}
}
\newglossaryentry{process}
{
  name=process,
  description={is a method in execution.  We do not make the distinction
      between processes and threads.  A process has a program counter, a register,
      and a stack}
}
\newglossaryentry{process variable}
{
  name=process variable,
  description={is a variable that is private to a single process and stored in
        its register}
}
\newglossaryentry{property}
{
  name=property,
  description={describes a set of execution traces or histories that are allowed by
    a program.  Safety properties are properties in which ``no bad things happen,''
    such as violating mutual exclusion in a critical section.  Liveness properties are
    properties where ``something good eventually happens,'' like processes being
    able to enter the critical section if they want to}
}
\newglossaryentry{sequential execution}
{
  name=sequential execution,
  description={is when there is just one process executing, as opposed to concurrent
    execution}
}
\newglossaryentry{shared variable}
{
  name=shared variable,
  description={is a variable that is stored in the memory of the CXL machine and
      shared between multiple processes, as opposed to a process variable}
}
\newglossaryentry{stack machine}
{
  name=stack machine,
  description={is a model of computing where the state of a process is kept on
    a stack.  CXL uses a combination of a stack machine and a register-based machine}
}
\newglossaryentry{thread safety}
{
  name=thread safety,
  description={is when the implementation of a data structure allows concurrent access
    with well-defined semantics}
}

@@

The result in this particular interleaving is that \texttt{count} ends up
being 1.  When running CXL, it will
report violations of assertions.  It also provides an example
of an interleaving, like the one above, in which an assertion fails.

>>>>

\newglossaryentry{atomicity}
{
  name=atomicity,
  description={describes that a certain machine instruction or sequence of machine
    instructions by a process is executed indivisibly and cannot be interleaved with 
    machine instructions of another process}
}
\newglossaryentry{concurrent execution}
{
  name=concurrent execution,
  description={(aka parallel execution) is when there are multiple processes executing and
      their machine instructions are interleaved in an unpredictable manner}
}
\newglossaryentry{context}
{
  name=context,
  description={(aka continuation) describes the state of a running process,
    including its program counter, the values of its variables (stored in its
    register), and the contents of its stack}
}
\newglossaryentry{critical section}
{
  name=critical section,
  description={(aka critical region) is a set of instructions that only one process
    is allowed to execute at a time.  The instructions are, however, not executed
    atomically, as other processes can continue to execute and access shared
    variables}
}
\newglossaryentry{determinism}
{
  name=determinism,
  description={is when the outcome of an execution is uniquely determined by the
    initial state}
}
\newglossaryentry{invariant}
{
  name=invariant,
  description={is a binary predicate over states that must hold for every
  reachable state of a process}
}
\newglossaryentry{machine instruction}
{
  name=machine instruction,
  description={is an atomic operation on the CXL machine, executed by a process}
}
\newglossaryentry{macro step}
{
  name=macro step,
  description={is a sequence of micro steps with at most one instruction
        externally visible to other processes.  It is the unit of interleaving
        in CXL}
}
\newglossaryentry{micro step}
{
  name=micro step,
  description={is the execution of one machine instruction by a process, updating
        its state and the state of the CXL machine}
}
\newglossaryentry{model checking}
{
  name=model checking,
  description={is a formal verification method that explores all possible executions
    of a program, which must have a finite number of states}
}
\newglossaryentry{process}
{
  name=process,
  description={is a method in execution.  We do not make the distinction
      between processes and threads.  A process has a program counter, a register,
      and a stack}
}
\newglossaryentry{process variable}
{
  name=process variable,
  description={is a variable that is private to a single process and stored in
        its register}
}
\newglossaryentry{property}
{
  name=property,
  description={describes a set of execution traces or histories that are allowed by
    a program.  Safety properties are properties in which ``no bad things happen,''
    such as violating mutual exclusion in a critical section.  Liveness properties are
    properties where ``something good eventually happens,'' like processes being
    able to enter the critical section if they want to}
}
\newglossaryentry{race condition}
{
  name=race condition,
  description={describes when multiple processes access shared state concurrently,
    leading to undesirable outcomes}
}
\newglossaryentry{sequential execution}
{
  name=sequential execution,
  description={is when there is just one process executing, as opposed to concurrent
    execution}
}
\newglossaryentry{shared variable}
{
  name=shared variable,
  description={is a variable that is stored in the memory of the CXL machine and
      shared between multiple processes, as opposed to a process variable}
}
\newglossaryentry{stack machine}
{
  name=stack machine,
  description={is a model of computing where the state of a process is kept on
    a stack.  CXL uses a combination of a stack machine and a register-based machine}
}
\newglossaryentry{thread safety}
{
  name=thread safety,
  description={is when the implementation of a data structure allows concurrent access
    with well-defined semantics}
}

@@

The result in this particular interleaving is that \texttt{count} ends up
being 1.
This is known as a \emph{race condition}.
\index{race condition}
\glsadd{race condition}
When running CXL, it will
report violations of assertions.  It also provides an example
of an interleaving, like the one above, in which an assertion fails.
